---
title: "Simulations and calculations of Linezolid PKPD index from LCS concentrations"
author: Michael BURTIN
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

#############################################################################
###########                       Depencies                       ###########         
#############################################################################
Load necessary library and depencies for the project. 
You can also specify different parameters as PKPD parameters, number of patients, etc.

```{r}
#| echo: false
#| include: false

####
## Load depencies
####
library(mrgsolve)       # Simulation package
library(dplyr)          # Data manipulation
library(tidyr)          # Data manipulation
library(purrr)          # Data manipulation
library(tibble)         # Data manipulation
library(ggplot2)        # Plotting

# If doesn't exist, create a folder for plots
if (!dir.exists("results")) {
  dir.create("results")
}

# Load the model
model_file <- mread("model/Linezolid_PKPD_model")

####
## Specify your entries
####

# Multithreading setup (define the max number to use for simulations)
thread_number = 9

# Global PKPD parameters
CL = 8.38
V1 = 36.4
FU = 0.69

# Specific parameters for each strains
model_list <- list(
    # S. Aureus models
    SA07982 = param(model_file, PCL=CL, B0 = 5.98, Bmax = 9.76, Kg = 1.04,
                   Emax = 2.16, EC50 = 1.42, Kon = 0.0, MIC = 1.0) |>
              smat(matrix(0.25)),
    SA29213 = param(model_file, PCL=CL, B0 = 6.02, Bmax = 9.35, Kg = 1.26,
                    Emax = 2.41, EC50 = 2.28, Kon = 0.014, MIC = 4.0) |> 
              smat(matrix(0.38)),
    SA01943 = param(model_file, PCL=CL, B0 = 6.09, Bmax = 9.26, Kg = 1.09,
                    Emax = 1.93, EC50 = 1.1, Kon = 0.016, MIC = 2.0) |>
              smat(matrix(0.35)),
    SA33591 = param(model_file, PCL=CL, B0 = 5.96, Bmax = 9.46, Kg = 1.02,
                    Emax = 1.52, EC50 = 0.68, Kon = 0.0, MIC = 1.0) |>
              smat(matrix(0.37)),
    # S. Epidermidis models
    SE17036 = param(model_file, PCL=CL, B0 = 5.81, Bmax = 9, Kg = 0.97,
                   Emax = 1.65, EC50 = 0.51, Kon = 0.021, MIC = 2.0) |>
              smat(matrix(0.35)),
    SE18436 = param(model_file, PCL=CL, B0 = 6.28, Bmax = 9.13, Kg = 0.86,
                   Emax = 1.11, EC50 = 0.54, Kon = 0.0, MIC = 2.0) |>
              smat(matrix(0.24)),
    SE14990 = param(model_file, PCL=CL, B0 = 5.77, Bmax = 8.98, Kg = 1.07,
                    Emax = 1.54, EC50 = 1.06, Kon = 0.0, MIC = 2.0) |>
              smat(matrix(0.64)),
    # S. Capitis models
    SC14982 = param(model_file, PCL=CL, B0 = 6.05, Bmax = 9.3, Kg = 0.83,
                   Emax = 1.56, EC50 = 1.34, Kon = 0.0, MIC = 2.0) |>
              smat(matrix(0.47)),
    SC19668 = param(model_file, PCL=CL, B0 = 6.38, Bmax = 9.3, Kg = 1.03,
                   Emax = 1.5, EC50 = 0.62, Kon = 0.0, MIC = 2.0) |>
              smat(matrix(0.42))
)

# Specify AMT choosen for fractioned doses
fractioned_daily_AMT <- c(10, 25, 50, 75, 100, 150, 300, 450, 600, 800, 1000, 1200, 1800, 2400, 3000, 4000, 5500, 7000, 8500, 10000)

# Specify the number of simulated patients
i_data <- expand.idata(ID = 1:250)

####
## Execute the main script
####
source("RScripts/main_script.R")

```

#############################################################################
###########               PKPD Plots generation                   ###########
#############################################################################
Generate PKPD plots binding Log10 CFU variation at 24h to all possible
PKPD indexes values. Five indexes are used in this study: fCmax/MIC, fAUC/MIC, 
fT>MIC, fT>MIC 4X, fT>MIC 10X.

```{r}
# Load the generic function to generate plots
source("RScripts/pkpd_plots_script.R")

# Generate the plots
csf_plot <- generate_PKPD_plots(obs_data = csf_data[["obs_mean"]], 
                                pred_data = csf_data[["pred_data"]], 
                                title = "LNZ PKPD Index for CSF" ,
                                filename = "LNZ PKPD Indexes for CSF target.png")

plasma_plot <- generate_PKPD_plots(obs_data = plasma_data[["obs_mean"]], 
                                   pred_data = plasma_data[["pred_data"]], 
                                   title = "LNZ PKPD Index for Plasma",
                                   "LNZ PKPD Indexes for Plasma target.png")

mix_plot <- generate_PKPD_plots(obs_data = csf_with_plasma_index_data[["obs_mean"]], 
                                pred_data = csf_with_plasma_index_data[["pred_data"]], 
                                title = "LNZ PKPD Index for CSF with Plasma indexes",
                                filename = "LNZ PKPD Indexes for CSF target with Plasma Indexes.png")

```

#############################################################################
###########          Target estimations for each indexes          ###########
#############################################################################
Generate a table for each PKPD index with index values for different PD targets 
(bacteriostasis, 1-log kill, 2-log kill).

```{r}
# Generic function to find the index value for a specific target. Based on B0 and pred data.
target_index_efficacity <- list(
  
csf_efficacity = csf_data[["correlation_data"]] |>
  select(c(
    PKPD_Index,
    Target_stasis,
    Target_1LogKill,
    Target_2LogKill,
  )) |>
  pivot_longer(-c(PKPD_Index),
               names_to = "EfficacyCriteria",
               values_to = "TargetIndexValue") |>
  mutate(TargetIndexValue =
           map_chr(TargetIndexValue,
                   ~ if(anyNA(.x)) {return(NA)
                   }
                   else
                   {
                     paste0(
                       round(.x[1, 1], 2), " [", round(.x[1, 3], 2), " - ", round(.x[1, 4], 2), "]"
                     )})) |>
  pivot_wider(names_from = PKPD_Index, values_from = TargetIndexValue),

plasma_efficacity = plasma_data[["correlation_data"]] |>
  select(c(
    PKPD_Index,
    Target_stasis,
    Target_1LogKill,
    Target_2LogKill,
  )) |>
  pivot_longer(-c(PKPD_Index),
               names_to = "EfficacyCriteria",
               values_to = "TargetIndexValue") |>
  mutate(TargetIndexValue =
           map_chr(TargetIndexValue,
                   ~ if(anyNA(.x)) {return(NA)
                   }
                   else
                   {
                     paste0(
                       round(.x[1, 1], 2), " [", round(.x[1, 3], 2), " - ", round(.x[1, 4], 2), "]"
                     )})) |>
  pivot_wider(names_from = PKPD_Index, values_from = TargetIndexValue),

mix_efficacity = csf_with_plasma_index_data[["correlation_data"]] |>
  select(c(
    PKPD_Index,
    Target_stasis,
    Target_1LogKill,
    Target_2LogKill,
  )) |>
  pivot_longer(-c(PKPD_Index),
               names_to = "EfficacyCriteria",
               values_to = "TargetIndexValue") |>
  mutate(TargetIndexValue =
           map_chr(TargetIndexValue,
                   ~ if(anyNA(.x)) {return(NA)
                   }
                   else
                   {
                     paste0(
                       round(.x[1, 1], 2), " [", round(.x[1, 3], 2), " - ", round(.x[1, 4], 2), "]"
                     )})) |>
  pivot_wider(names_from = PKPD_Index, values_from = TargetIndexValue)
)

```

#############################################################################
###########                 Intermediate analysis                 ###########
#############################################################################
Various analysis that can resume our data.

```{r}
# Resume several differents data in a list
summary_data <- list(
# Resume fractioned data at 24h for each AMT & Dose Group
fractioned_24h = sim_results_fractioned[["fractioned_results"]][[1]] |>
    group_by(DoseGroup, AMT) |>
    summarize( 
      C_CENTRAL_mean = mean(C_CENTRAL),
      C_CENTRAL_by_AUC = mean(AUCCENTRAL[time == 24.0])/24,
      C_CSF_mean = mean(C_CSF),
      C_CSF_by_AUC = mean(AUCCSF[time == 24.0])/24,
      Cmax_CENTRAL_mean = mean(Cmax_CENTRAL[time == 24.0]),
      Cmax_CSF = mean(Cmax_CSF[time == 24.0])
    ),

# Resume fractioned data at 24h for each AMT
fractioned_24h_by_AMT = sim_results_fractioned[["fractioned_results"]][[1]] |>
    group_by(AMT) |>
    summarize( 
      C_CENTRAL_mean = mean(C_CENTRAL),
      C_CENTRAL_by_AUC = mean(AUCCENTRAL[time == 24.0])/24,
      C_CSF_mean = mean(C_CSF),
      C_CSF_by_AUC = mean(AUCCSF[time == 24.0])/24,
      Cmax_CENTRAL_mean = mean(Cmax_CENTRAL[time == 24.0]),
      Cmax_CSF = mean(Cmax_CSF[time == 24.0])
    ),

# Resume all fractioned data over the time for PD & PK
fractioned_full = sim_results_fractioned[["fractioned_results"]][[1]] |>
    group_by(DoseGroup, AMT, time) |>
    summarize( 
      Log10CFU_CENTRAL = mean(Log10CFU_CENTRAL),
      Log10CFU_CSF = mean(Log10CFU_CSF),
      C_CENTRAL = mean(C_CENTRAL),
      C_CSF = mean(C_CSF)
    ),

# Resume all fractioned data over the time for PD & PK
continuous_full = sim_results_continuous_inf[["continuous_inf_results"]] |>
    group_by(STRN, AMT, time) |>
    summarize( 
      Log10CFU_CENTRAL = mean(Log10CFU_CENTRAL),
      Log10CFU_CSF = mean(Log10CFU_CSF),
      C_CENTRAL = mean(C_CENTRAL),
      C_CSF = mean(C_CSF)
    )
)

# Resume the percentage of data under a certain log-target
targets_summary <- list(
  
  csf_global = csf_data[["obs_mean"]] |>
  ungroup() |>
  summarise(
    Stasis_reach = mean(deltaLog10CFU > 0) * 100,
    Log1Kill_reach = mean(deltaLog10CFU < -1) * 100,
    Log2Kill_reach = mean(deltaLog10CFU < -2) * 100
  ),
  
  csf_by_DoseGroup = csf_data[["obs_mean"]] |>
  ungroup() |>
  group_by(DoseGroup) |>
  summarise(
    Stasis_reach = mean(deltaLog10CFU < 0) * 100,
    Log1Kill_reach = mean(deltaLog10CFU < -1) * 100,
    Log2Kill_reach = mean(deltaLog10CFU < -2) * 100,
  ) |> filter(DoseGroup != "control"),
  
  csf_by_AMT = csf_data[["obs_mean"]] |>
  ungroup() |>
  group_by(AMT) |>
  summarise(
    Stasis_reach = mean(deltaLog10CFU < 0) * 100,
    Log1Kill_reach = mean(deltaLog10CFU < -1) * 100,
    Log2Kill_reach = mean(deltaLog10CFU < -2) * 100,
  ),
  
  plasma_global = plasma_data[["obs_mean"]] |>
    ungroup() |>
    summarise(
      Stasis_reach = mean(deltaLog10CFU < 0) * 100,
      Log1Kill_reach = mean(deltaLog10CFU < -1) * 100,
      Log2Kill_reach = mean(deltaLog10CFU < -2) * 100
    ),
  
  plasma_by_DoseGroup = plasma_data[["obs_mean"]] |>
  ungroup() |>
  group_by(DoseGroup) |>
  summarise(
    Stasis_reach = mean(deltaLog10CFU < 0) * 100,
    Log1Kill_reach = mean(deltaLog10CFU < -1) * 100,
    Log2Kill_reach = mean(deltaLog10CFU < -2) * 100
  ) |> filter(DoseGroup != "control"),
  
  plasma_by_AMT = plasma_data[["obs_mean"]] |>
  ungroup() |>
  group_by(AMT) |>
  summarise(
    Stasis_reach = mean(deltaLog10CFU < 0) * 100,
    Log1Kill_reach = mean(deltaLog10CFU < -1) * 100,
    Log2Kill_reach = mean(deltaLog10CFU < -2) * 100
  )
)

```

#############################################################################
###########                      VPC generation                   ###########
#############################################################################
Generate all VPC objects needed for plots.

```{r}
source("RScripts/vpc_script.R")
library(tidyvpc)

# For T>MIC
vpc_by_DoseGroup_ToverMIC_plasma <- observed(obs_plasma_data |> filter(PKPD_Index == "I3_ToverMIC"), x = Index_value, y = DV) |>
  simulated(isim_plasma_data, y = IPRED) |>
  stratify(~ DoseGroup) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

vpc_global_ToverMIC_plasma <- observed(obs_plasma_data |> filter(PKPD_Index == "I3_ToverMIC"), x = Index_value, y = DV) |>
  simulated(isim_plasma_data, y = IPRED) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

vpc_by_DoseGroup_ToverMIC_csf <- observed(obs_csf_data |> filter(PKPD_Index == "I3_ToverMIC"), x = Index_value, y = DV) |>
  simulated(isim_plasma_data, y = IPRED) |>
  stratify(~ DoseGroup) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

vpc_global_ToverMIC_csf <- observed(obs_csf_data |> filter(PKPD_Index == "I3_ToverMIC"), x = Index_value, y = DV) |>
  simulated(isim_plasma_data, y = IPRED) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

# For AUC
vpc_by_DoseGroup_AUC_plasma <- observed(obs_plasma_data |> filter(PKPD_Index == "I2_AUC"), x = Index_value, y = DV) |>
  simulated(isim_plasma_data, y = IPRED) |>
  stratify(~ DoseGroup) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

vpc_global_AUC_plasma <- observed(obs_plasma_data |> filter(PKPD_Index == "I2_AUC"), x = Index_value, y = DV) |>
  simulated(isim_plasma_data, y = IPRED) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

vpc_by_DoseGroup_AUC_csf <- observed(obs_csf_data |> filter(PKPD_Index == "I2_AUC"), x = Index_value, y = DV) |>
  simulated(isim_csf_data, y = IPRED) |>
  stratify(~ DoseGroup) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

vpc_global_AUC_csf <- observed(obs_csf_data |> filter(PKPD_Index == "I2_AUC"), x = Index_value, y = DV) |>
  simulated(isim_csf_data, y = IPRED) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

# Cmax
vpc_by_DoseGroup_Cmax_plasma <- observed(obs_plasma_data |> filter(PKPD_Index == "I1_Cmax"), x = Index_value, y = DV) |>
  simulated(isim_plasma_data, y = IPRED) |>
  stratify(~ DoseGroup) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

vpc_global_Cmax_plasma <- observed(obs_plasma_data |> filter(PKPD_Index == "I1_Cmax"), x = Index_value, y = DV) |>
  simulated(isim_plasma_data, y = IPRED) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

vpc_by_DoseGroup_Cmax_csf <- observed(obs_csf_data |> filter(PKPD_Index == "I1_Cmax"), x = Index_value, y = DV) |>
  simulated(isim_csf_data, y = IPRED) |>
  stratify(~ DoseGroup) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()

vpc_global_Cmax_csf <- observed(obs_csf_data |> filter(PKPD_Index == "I1_Cmax"), x = Index_value, y = DV) |>
  simulated(isim_csf_data, y = IPRED) |>
  binning(bin = "ntile", nbins = 9) |>
  vpcstats()
```

Generate all VPC plots. Need to execute the block above.

```{r}
## Plotting Plasma
plot(vpc_by_DoseGroup_ToverMIC_plasma, log = "x", show.points = FALSE) + 
  scale_x_continuous(limits = c(0, 100), breaks = c(0, 25, 50, 75, 100)) + labs(x = "fT>MIC (%)", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for T>MIC in Plasma by Dose Group")
ggsave("results/VPC_ToverMIC_plasma_by_DoseGroup.png")

plot(vpc_global_ToverMIC_plasma, log = "x", show.points = FALSE) + 
  scale_x_continuous(limits = c(0, 100), breaks = c(0, 25, 50, 75, 100)) + labs(x = "fT>MIC (%)", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for T>MIC in Plasma")
ggsave("results/VPC_ToverMIC_plasma.png")

plot(vpc_by_DoseGroup_AUC_plasma, log = "x", show.points = FALSE) + 
  scale_x_continuous(limits = c(1, 1000), breaks = c(1, 10, 100, 1000), trans="log10") + labs(x = "fAUC/MIC", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for AUC in Plasma by Dose Group")
ggsave("results/VPC_AUC_plasma_by_DoseGroup.png")

plot(vpc_global_AUC_plasma, log = "x", show.points = FALSE) + 
  scale_x_continuous(limits = c(1, 1000), breaks = c(1, 10, 100, 1000), trans="log10") + labs(x = "fAUC/MIC", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for AUC in Plasma")
ggsave("results/VPC_AUC_plasma.png")

plot(vpc_by_DoseGroup_Cmax_plasma, log = "x", show.points = FALSE) + 
  scale_x_continuous(limits = c(0.1, 100), breaks = c(0.1, 1, 10, 100), trans="log10") + labs(x = "fCmax/MIC (mg/l)", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for Cmax in Plasma by Dose Group")
ggsave("results/VPC_Cmax_plasma_by_DoseGroup.png")

plot(vpc_global_Cmax_plasma, log = "x", show.points = FALSE) + 
  scale_x_continuous(limits = c(0.1, 100), breaks = c(0.1, 1, 10, 100), trans="log10") + labs(x = "fCmax/MIC (mg/l)", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for Cmax in Plasma")
ggsave("results/VPC_Cmax_plasma.png")

## Plotting CSF
plot(vpc_by_DoseGroup_ToverMIC_csf, log = "x", show.points = FALSE) + scale_x_continuous(limits = c(0, 100), breaks = c(0, 25, 50, 75, 100)) + labs(x = "fT>MIC (%)", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for T>MIC in CSF by Dose Group")
ggsave("results/VPC_ToverMIC_csf_by_DoseGroup.png")

plot(vpc_global_ToverMIC_csf, log = "x", show.points = FALSE) + scale_x_continuous(limits = c(0, 100), breaks = c(0, 25, 50, 75, 100)) + labs(x = "fT>MIC (%)", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for T>MIC in CSF")
ggsave("results/VPC_ToverMIC_csf.png")

plot(vpc_by_DoseGroup_AUC_csf, log = "x", show.points = FALSE) + scale_x_continuous(limits = c(1, 1000), breaks = c(1, 10, 100, 1000), trans="log10") + labs(x = "fAUC/MIC", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for AUC in CSF by Dose Group")

ggsave("results/VPC_AUC_csf_by_DoseGroup.png")
plot(vpc_global_AUC_csf, log = "x", show.points = FALSE) + scale_x_continuous(limits = c(1, 1000), breaks = c(1, 10, 100, 1000), trans="log10") + labs(x = "fAUC/MIC", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for AUC in CSF")
ggsave("results/VPC_AUC_csf.png")

plot(vpc_by_DoseGroup_Cmax_csf, log = "x", show.points = FALSE) + scale_x_continuous(limits = c(0.1, 50), breaks = c(0.1, 1, 10, 50), trans="log10") + labs(x = "fCmax/MIC (mg/l)", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for Cmax in CSF by Dose Group")
ggsave("results/VPC_Cmax_csf_by_DoseGroup.png")

plot(vpc_global_Cmax_csf, log = "x", show.points = FALSE) + scale_x_continuous(limits = c(0.1, 50), breaks = c(0.1, 1, 10, 50), trans="log10") + labs(x = "fCmax/MIC (mg/l)", y = "deltaLog10CFU (CFU/ml)") + ggtitle("VPC for Cmax in CSF")
ggsave("results/VPC_Cmax_csf.png")
```

#############################################################################
###########                  Stratification plots                ###########
#############################################################################
```{r}
ggplot(filter(csf_data[["obs_mean"]], PKPD_Index == "I2_AUC"))  +
      geom_point(mapping = aes(
        x = value,
        y = deltaLog10CFU,
        color = as.factor(DoseGroup)
      )) +
    scale_x_continuous(
      limits = c(1, 1000),
      breaks = c(1, 10, 100, 1000), 
      trans="log10"
    ) +
    geom_line(filter(csf_data[["pred_data"]], PKPD_Index == "I2_AUC"), 
              mapping = aes(x=value, y=pred), linewidth = 0.8) +
    geom_text(aes(x = Inf, y = Inf, label = paste("R² = ", round(Adj.Rsq, 2))), 
              hjust = 1.2, vjust = 2, size = 4, fontface = "italic", color = "black") +
  theme_bw()

ggplot(filter(csf_data[["obs_mean"]], PKPD_Index == "I2_AUC"))  +
      geom_point(mapping = aes(
        x = value,
        y = deltaLog10CFU,
        color = as.factor(AMT)
      )) +
    scale_x_continuous(
      limits = c(1, 1000),
      breaks = c(1, 10, 100, 1000), 
      trans="log10"
    ) +
    geom_line(filter(csf_data[["pred_data"]], PKPD_Index == "I2_AUC"), 
              mapping = aes(x=value, y=pred), linewidth = 0.8) +
    geom_text(aes(x = Inf, y = Inf, label = paste("R² = ", round(Adj.Rsq, 2))), 
              hjust = 1.2, vjust = 2, size = 4, fontface = "italic", color = "black") +
  theme_bw()


ggplot(filter(csf_data[["obs_mean"]], PKPD_Index == "I3_ToverMIC"))  +
      geom_point(mapping = aes(
        x = value,
        y = deltaLog10CFU,
        color = as.factor(DoseGroup)
      )) +
    geom_line(filter(csf_data[["pred_data"]], PKPD_Index == "I3_ToverMIC"), 
              mapping = aes(x=value, y=pred), linewidth = 0.8) +
    geom_text(aes(x = Inf, y = Inf, label = paste("R² = ", round(Adj.Rsq, 2))), 
              hjust = 1.2, vjust = 2, size = 4, fontface = "italic", color = "black") +
  theme_bw()

ggplot(filter(csf_data[["obs_mean"]], PKPD_Index == "I3_ToverMIC"))  +
      geom_point(mapping = aes(
        x = value,
        y = deltaLog10CFU,
        color = as.factor(AMT)
      )) +
    geom_line(filter(csf_data[["pred_data"]], PKPD_Index == "I3_ToverMIC"), 
              mapping = aes(x=value, y=pred), linewidth = 0.8) +
    geom_text(aes(x = Inf, y = Inf, label = paste("R² = ", round(Adj.Rsq, 2))), 
              hjust = 1.2, vjust = 2, size = 4, fontface = "italic", color = "black") +
  theme_bw()

ggplot(filter(csf_data[["obs_mean"]], PKPD_Index == "I4_ToverMIC_4X"))  +
      geom_point(mapping = aes(
        x = value,
        y = deltaLog10CFU,
        color = as.factor(DoseGroup)
      )) +
    geom_line(filter(csf_data[["pred_data"]], PKPD_Index == "I4_ToverMIC_4X"), 
              mapping = aes(x=value, y=pred), linewidth = 0.8) +
    geom_text(aes(x = Inf, y = Inf, label = paste("R² = ", round(Adj.Rsq, 2))), 
              hjust = 1.2, vjust = 2, size = 4, fontface = "italic", color = "black") +
  theme_bw()

ggplot(filter(csf_data[["obs_mean"]], PKPD_Index == "I4_ToverMIC_4X"))  +
      geom_point(mapping = aes(
        x = value,
        y = deltaLog10CFU,
        color = as.factor(AMT)
      )) +
    geom_line(filter(csf_data[["pred_data"]], PKPD_Index == "I4_ToverMIC_4X"), 
              mapping = aes(x=value, y=pred), linewidth = 0.8) +
    geom_text(aes(x = Inf, y = Inf, label = paste("R² = ", round(Adj.Rsq, 2))), 
              hjust = 1.2, vjust = 2, size = 4, fontface = "italic", color = "black") +
  theme_bw()

ggplot(filter(csf_data[["obs_mean"]], PKPD_Index == "I5_ToverMIC_10X"))  +
      geom_point(mapping = aes(
        x = value,
        y = deltaLog10CFU,
        color = as.factor(DoseGroup)
      )) +
    geom_line(filter(csf_data[["pred_data"]], PKPD_Index == "I5_ToverMIC_10X"), 
              mapping = aes(x=value, y=pred), linewidth = 0.8) +
    geom_text(aes(x = Inf, y = Inf, label = paste("R² = ", round(Adj.Rsq, 2))), 
              hjust = 1.2, vjust = 2, size = 4, fontface = "italic", color = "black") +
  theme_bw()

ggplot(filter(csf_data[["obs_mean"]], PKPD_Index == "I5_ToverMIC_10X", AMT >= 1200))  +
      geom_point(mapping = aes(
        x = value,
        y = deltaLog10CFU,
        color = as.factor(AMT)
      )) +
    geom_line(filter(csf_data[["pred_data"]], PKPD_Index == "I5_ToverMIC_10X"), 
              mapping = aes(x=value, y=pred), linewidth = 0.8) +
    geom_text(aes(x = Inf, y = Inf, label = paste("R² = ", round(Adj.Rsq, 2))), 
              hjust = 1.2, vjust = 2, size = 4, fontface = "italic", color = "black") +
  theme_bw()
```

#############################################################################
###########                       PK Plots                        ###########
#############################################################################

```{r}
# Plasma plot of LNZ conc across time by AMT with Continuous infusions
summary_data[["continuous_full"]] |>
  filter(STRN == "A07982") |>
  ggplot(aes(x = time, y = C_CENTRAL)) +
  geom_line() +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "Plasmatic PK of Linezolid for each dosing regimens",
    color = "Dose amount (mg)"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/plasma_pk_continuous_infusions_by_AMT.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

# Plasma plot of LNZ conc across time by AMT
summary_data[["fractioned_full"]] |>
  filter(AMT != 0) |>
  ggplot(aes(x = time, y = C_CENTRAL, color = as.factor(AMT))) +
  geom_line() +
  facet_wrap(~DoseGroup) +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "Plasmatic PK of Linezolid for each dosing regimens",
    color = "Dose amount (mg)"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/plasma_pk_by_AMT.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

# Plasma plot of LNZ conc across time by DoseGroup
summary_data[["fractioned_full"]] |>
  ggplot(aes(x = time, y = C_CENTRAL, color = as.factor(DoseGroup))) +
  geom_line() +
  facet_wrap(~AMT, ncol = 3) +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "Plasmatic PK of Linezolid for each dose amount",
    color = "Dose regimens"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/plasma_pk_by_DoseGroup.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

# Plasma plot of LNZ conc across time by DoseGroup with free y-scaling
summary_data[["fractioned_full"]] |>
  filter(AMT != 0) |>
  ggplot(aes(x = time, y = C_CENTRAL, color = as.factor(DoseGroup))) +
  geom_line() +
  facet_wrap(~AMT, ncol = 3, scales = "free_y") +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "Plasmatic PK of Linezolid for each dose amount",
    color = "Dose regimens"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

# Save plot with size A4
ggsave("results/plasma_pk_by_DoseGroup_free_y.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

# Plasma plot of LNZ conc across time by AMT
summary_data[["fractioned_full"]] |>
  ggplot(aes(x = time, y = C_CSF, color = as.factor(AMT))) +
  geom_line() +
  facet_wrap(~DoseGroup) +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "CSF PK of Linezolid for each dosing regimens",
    color = "Dose amount (mg)"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/csf_pk_by_AMT.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

# Plasma plot of LNZ conc across time by DoseGroup
summary_data[["fractioned_full"]] |>
  filter(AMT != 0) |>
  ggplot(aes(x = time, y = C_CSF, color = as.factor(DoseGroup))) +
  geom_line() +
  facet_wrap(~AMT, ncol = 3) +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "CSF PK of Linezolid for each dose amount",
    color = "Dose regimens"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/csf_pk_by_DoseGroup.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

# Plasma plot of LNZ conc across time by DoseGroup with free y-scaling
summary_data[["fractioned_full"]] |>
  filter(AMT != 0) |>
  ggplot(aes(x = time, y = C_CSF, color = as.factor(DoseGroup))) +
  geom_line() +
  facet_wrap(~AMT, ncol = 3, scales = "free_y") +  # Échelle 'y' libre pour chaque facette
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "CSF PK of Linezolid for each dose amount",
    color = "Dose regimens"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/csf_pk_by_DoseGroup_free_y.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

```

#############################################################################
###########                       PD Plots                        ###########
#############################################################################

```{r}
summary_full |>
  ggplot(aes(x = time, y = Log10CFU_CENTRAL, color = as.factor(DoseGroup))) +
  geom_line() +
  facet_wrap(~AMT, ncol = 3) +  # Échelle 'y' libre pour chaque facette
  labs(
    x = "Time (h)", 
    y = "Number of bacteria in (log10, CFU/ml) (mg/L)", 
    title = "Plasmatic PD of A0798 for each dose amount of Linezolid",
    color = "Dose regimens"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/plasma_pd_by_AMT.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

summary_full |>
  ggplot(aes(x = time, y = Log10CFU_CENTRAL, color = as.factor(AMT))) +
  geom_line() +
  facet_wrap(~DoseGroup) +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "Plasmatic PD of A07982 for each dosing regimens of Linezolid",
    color = "Dose amount (mg)"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/plasma_pd_by_DoseGroup.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

summary_full |>
  ggplot(aes(x = time, y = Log10CFU_CSF, color = as.factor(DoseGroup))) +
  geom_line() +
  facet_wrap(~AMT, ncol = 3) +  # Échelle 'y' libre pour chaque facette
  labs(
    x = "Time (h)", 
    y = "Number of bacteria in (log10, CFU/ml) (mg/L)", 
    title = "CSF PD of A0798 for each dose amount of Linezolid",
    color = "Dose regimens"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/csf_pd_by_AMT.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

summary_full |>
  ggplot(aes(x = time, y = Log10CFU_CSF, color = as.factor(AMT))) +
  geom_line() +
  facet_wrap(~DoseGroup) +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "CSF PD of A0798 for each dosing regimens of Linezolid",
    color = "Dose amount (mg)"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))

ggsave("results/csf_pd_by_DoseGroup.png", width = 8.27, height = 11.69, units = "in", dpi = 300)

```
