---
title: "Study of PK/PD Index for Linezolid specific to CSF infections"
editor_options: 
  chunk_output_type: console
---

## Settings & Dependencies

Load necessary library, dependencies and settings for the project. Just run the block code to generate the same result than the article. Otherwise, if you want to play simulations, you can change the parameters, dosing regimens...

```{r}
#| echo: false

####
## Load dependencies
####
library(mrgsolve)       # Simulation package
library(dplyr)          # Data manipulation
library(tidyr)          # Data manipulation
library(purrr)          # Data manipulation
library(tibble)         # Data manipulation
library(ggplot2)        # Plotting

# If doesn't exist, create a folder for result plots
if (!dir.exists("results")) {
  dir.create("results")
}

# Load the model
model_file <- mread("model/Mouse_model_v3")

####
## Specify your entries
####

# Define the maximum thread number to use for simulations
thread_number = 9

# These PK parameters will be used later in the Rscripts.
# So we define them here, to avoid to hardcode values.
V1 = 1.31

#    SA07982 = param(model_file, B0 = 5.98, Bmax = 9.76, Kg = 1.04,
#                   Emax = 2.16, EC50 = 1.42, Kon = 0.0, MIC = 1.0),
#    SA29213 = param(model_file, B0 = 6.02, Bmax = 9.35, Kg = 1.26,
#                    Emax = 2.41, EC50 = 2.28, Kon = 0.014, MIC = 4.0),
#    SA01943 = param(model_file, B0 = 6.09, Bmax = 9.26, Kg = 1.09,
#                    Emax = 1.93, EC50 = 1.1, Kon = 0.016, MIC = 2.0),
#    SA33591 = param(model_file, B0 = 5.96, Bmax = 9.46, Kg = 1.02,
#                    Emax = 1.52, EC50 = 0.68, Kon = 0.0, MIC = 1.0),

# Specific model parameters for each strains
model_list <- list(
    # S. Aureus
    SA29213 = param(model_file, B0 = 6.02, Bmax = 9.35, Kg = 1.26,
                   Emax = 2.41, EC50 = 2.28, Kon = 0.014, MIC = 4.0)
)

dose0 <- ev(ID = 0, amt = 0, ii= 24, addl = 0, cmt ="A")
dose1 <- ev(ID = 1, amt = 72, ii= 24, addl = 0, cmt ="A")
dose2 <- ev(ID = 2, amt = 80, ii= 12, addl = 1, cmt ="A")
dose3 <- ev(ID = 3, amt = 60, ii= 12, addl = 1, cmt ="A")
dose4 <- ev(ID = 4, amt = 44, ii= 12, addl = 1, cmt ="A")
dose5 <- ev(ID = 5, amt = 72, ii= 8, addl = 2, cmt ="A")
dose6 <- ev(ID = 6, amt = 80, ii= 6, addl = 3, cmt ="A")
dose7 <- ev(ID = 7, amt = 72, ii= 6, addl = 3, cmt ="A")
dose8 <- ev(ID = 8, amt = 60, ii= 6, addl = 3, cmt ="A")
dose9 <- ev(ID = 9, amt = 48, ii= 6, addl = 3, cmt ="A")
dose10 <- ev(ID = 10, amt = 60, ii= 4, addl = 5, cmt ="A")
dose11 <- ev(ID = 11, amt = 44, ii= 4, addl = 5, cmt ="A")
dose12 <- ev(ID = 12, amt = 36, ii= 4, addl = 5, cmt ="A")
dose13 <- ev(ID = 13, amt = 24, ii= 4, addl = 5, cmt ="A")
dose14 <- ev(ID = 14, amt = 16, ii= 4, addl = 5, cmt ="A")

subcut_dose = rbind(dose0@data, dose1@data, dose2@data, dose3@data, dose4@data, 
                    dose5@data, dose6@data, dose7@data, dose8@data,
                    dose9@data, dose10@data, dose11@data, dose12@data,
                    dose13@data, dose14@data)

rm(dose0, dose1, dose2, dose3, dose4, dose5, dose6, dose7, dose8,
   dose9, dose10, dose11, dose12, dose13, dose14)

####
## Execute the main script
####
source("RScripts/main_script_for_mouse.R")

```

## PKPD Plots generation

The code below will generate PK/PD index plots for fCmax/MIC, fAUC/MIC and fT>MIC for CSF target.

```{r}
source("RScripts/pkpd_index_plots_script.R")

pkpd_plot <- PKPD_index_plots(obs_data = data[["sim_data"]], 
                             pred_data = data[["pred_data"]], 
                             title = "Reproduction of Sandberg PKPD Index for Linezolid in the mouse" ,
                             file_name = "Sandberg PKPD Index for Linezolid.png")
```

# Differents PK Trials

```{r}
#| echo: false
#| include: false

####
## Load depencies
####
library(mrgsolve)       # Simulation package
library(tidyverse)

# If doesn't exist, create a folder for plots
if (!dir.exists("results")) {
  dir.create("results")
}

# Load all differents PK studies results
bigelow_data <- read_csv("data/bigelow_data_AIO.csv")
sandberg_data <- read_csv("data/sandberg_free_concentration_AIO.csv")
andes_data <- read_csv("data/andes_data_AIO.csv")
```

Bioavailibility analysis
```{r}
library(minpack.lm)

# Determine a Michael model for F
f_calculation_df <- tibble(f = c(0.184, 0.233, 1, 1),
                           dose = c(10, 30, 100, 335))

# fit nls model of f_calculation_df with f in y and dose in x, and a michaelis menten mode
bioavailibility_model_01 <- nlsLM(f ~ Vmax * dose / (Km + dose), data = f_calculation_df, start = list(Vmax = 1, Km = 1))

## Result: Vmax = 1.290687 & Km = 63.448978

# another fit nls with Vmax between [0;1]
bioavailibility_model_02 <- nlsLM(f ~ Vmax * dose / (Km + dose), 
                                data = f_calculation_df, 
                                start = list(Vmax = 0.5, Km = 1), 
                                lower = c(Vmax = 0, Km = 0), 
                                upper = c(Vmax = 1, Km = Inf))

# Result: Vmax = 1.0 & Km = 36.12112
```

# Bigelow PK
```{r}
# Load the model
mouse_model <- mread("model/Mouse_model_bigelow")

## Bigelow results (PK)
ev01 <- ev(ID = 335, amt = 335, cmt = "A")
ev02 <- ev(ID = 100, amt = 100, cmt = "A")
ev03 <- ev(ID = 30, amt = 30, cmt = "A")
ev04 <- ev(ID = 10, amt = 10, cmt = "A")

dose_list <- rbind(ev01@data, ev02@data, ev03@data, ev04@data)

mouse_sim_01 <- mouse_model |>
    mrgsolve::data_set(dose_list) |>
    mrgsolve::mrgsim(delta = 0.1, end = 24) |>
    tibble::as_tibble()

## GGplot of test with time in x and C_CENTRAL in y, and color by ID
ggplot() +
    geom_line(data = mouse_sim_01, aes(x = time, y = C_CENTRAL, color = factor(ID))) +
    geom_line(data = bigelow_data, aes(x = time, y = conc, color = factor(ID)), linetype = "dashed") +
    labs(title = "Plasmatic concentration of Linezolid in mouse",
         x = "Time (h)",
         y = "Concentration (mg/L)",
         color = "ID") +
    scale_color_manual(values = c("335" = "purple", "100" = "cyan", "30" = "green4", "10" = "orange"), 
                       labels = c("335" = "335 mg/kg", "100" = "100 mg/kg", "30" = "30 mg/kg", "10" = "10 mg/kg")) +
    scale_y_continuous(breaks = c(0.0001, 0.001, 0.01, 0.1, 1, 10, 100, 1000), limits = c(0.0001, 1000), trans = "log10", 
                       labels = c("0.0001", "0.001", "0.01", "0.1", "1", "10",  "100", "1000")) +
    scale_x_continuous(breaks = seq(0, 24, 4)) +
    theme_minimal()

```

# Andes PK
```{r}
mouse_model <- mread("model/Mouse_model_v3")

ev05 <- ev(ID = 80, amt = 80, cmt ="A")
ev06 <- ev(ID = 40, amt = 40, cmt = "A")
ev07 <- ev(ID = 20, amt = 20, cmt = "A")
subcut_dose <- rbind(ev05@data, ev06@data, ev07@data)

mouse_sim_02 <- mouse_model |>
    mrgsolve::data_set(subcut_dose) |>
    mrgsolve::mrgsim(delta = 0.1, end = 24) |>
    tibble::as_tibble()

ggplot() +
    geom_line(data = mouse_sim_02, aes(x = time, y = C_CENTRAL, color = factor(ID))) +
    geom_line(data = sandberg_data, aes(x = time, y = fconc, color = factor(ID)), linetype = "dashed") +
    geom_point(data = sandberg_data, aes(x = time, y = fconc, color = factor(ID)), shape = 12) +
    scale_y_continuous(limits = c(0, 80), breaks = seq(0, 80, 10)) +
    scale_x_continuous(limits = c(0, 3)) +
    labs(title = "Plasmatic concentration of Linezolid in mouse - Comparaison between PK simulation (plain lines) & Andes (square dots) / Sandberg (triangle dots) data",
         x = "Time (h)",
         y = "Concentration (mg/L)",
         color = "Dose") +
    theme_minimal()

```


## PK & PD Analysis

```{r}
## Summarize data
summary_fractioned <- sim_results_fractioned[["fractioned_results"]][["SA6538p"]] |>
  group_by(DoseGroup, AMT, time) |>
  summarize(
    C_CENTRAL = mean(C_CENTRAL),
    Log10CFU = mean(Log10CFU)
    )

## PK plots
### Plasma PK for fractioned doses
summary_fractioned |>
  filter(AMT != 0) |>
  ggplot(aes(x = time, y = C_CENTRAL, color = as.factor(DoseGroup))) +
  geom_line(linewidth = 0.75) +
  facet_wrap(~AMT, ncol = 3, scales = "free_y") +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "Plasma PK of Linezolid in the mouse",
    color = "Dosing regimens"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))
### PD for fractioned doses
summary_fractioned |>
  filter(AMT != 0) |>
  ggplot(aes(x = time, y = Log10CFU, color = as.factor(DoseGroup))) +
  geom_line(linewidth = 0.75) +
  facet_wrap(~AMT, ncol = 3, scales = "free_y") +
  labs(
    x = "Time (h)", 
    y = "Linezolid concentrations (mg/L)", 
    title = "PD of Linezolid in the mouse",
    color = "Dosing regimens"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4))
```
